The universal property of direct integrals in W$^*$-categories

### 2018-04-30 direct_integrals.tex
proof

\documentclass[reqno,T1]{amsproc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multicol}
\usepackage[canadian]{babel}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{xifthen} % provides \isempty
\usepackage{xparse} % new commands with more than one optional argument

@@

\begin{thm}
The following are equivalent:
\begin{enumerate}
\item\label{universalprop} $\bigoplus_i A_i$ is an $I$-indexed direct sum.
\item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$ such that $\kappa_j^*\kappa_k = \delta_{jk}$, and
\beq
\label{complete}
	\sum_j \kappa_j \kappa_j^* = 1_{\bigoplus_i A_i}
\eeq
ultraweakly.
\end{enumerate}
\end{thm}

@@

Assuming \ref{biprodsum}, \todo{write this}
\end{proof}

>>>>

\documentclass[reqno,T1]{amsproc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multicol}
\usepackage[canadian]{babel}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{xifthen} % provides \isempty
\usepackage{xparse} % new commands with more than one optional argument
\newcommand{\unlim}{\qopname\relax m{unlim}}

@@

\begin{thm}
The following are equivalent:
\begin{enumerate}
\item\label{universalprop} $\bigoplus_i A_i$ is an $I$-indexed direct sum.
\item\label{universalprop2} 
There is a family of
        morphisms~$\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$
        with~$\sum_{j \in I} \kappa_j \kappa_j^* < \infty$
    such that for any object~$B$
        and family of
        morphisms~$\left(T_j: A_j \to B\right)_{j\in I}$
        with~$\sum_{j \in I}T_jT_j^* < \infty$,
        there is a unique~$T\colon \bigoplus_i A_i \to B$
        with~$T \kappa_j = T_j$ for all~$j \in I$.
\item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$ such that $\kappa_j^*\kappa_k = \delta_{jk}$, and
\beq
\label{complete}
        \sum_{j\in I} \kappa_j \kappa_j^* \ =\  1_{\bigoplus_i A_i}.
\eeq
\end{enumerate}
\end{thm}

@@

To finish, we will prove show \ref{biprodsum} $\Rightarrow$
        \ref{universalprop2}.
Let~$\left(T_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
        with~$\sum_{j \in I}T_jT_j^* < \infty$.
The homset~$C(B,A)$
    is a self-dual Hilbert~$C(B,B)$-module
    with~$C(B,B)$-valued inner
    product~$\langle S, T\rangle \equiv S^*T$.\cite[2.15]{wstarcat}
The self-duality is equivalent to the fact
    that~$C(B,A)$ is complete in the \emph{ultranorm} topology;
    that is, the uniformity
    generated by the seminorms~$\| T \|_\omega \equiv \omega(\langle T,T\rangle)^\frac{1}{2}$
    for normal states~$\omega$ on~$C(B,B)$.\cite[\S149\textsubscript{V}]{bas}
    Pick any normal state~$\omega$ on~$C(B,B)$.
        and finite subset~$S \subseteq I$.  We have
\begin{alignat*}{2}
    \bigl\| \sum_{j \in S} \kappa_j T_j^* \bigr\|^2_\omega
    &\ \equiv \ 
    \omega\Bigl(\bigl(\sum_{i \in S} \kappa_i T_i^*\bigr)^*
            \bigl(\sum_{j \in S} \kappa_j T_j^*\bigr)\Bigr) \\
    &\ =\ 
    \omega\Bigl( \sum_{j \in S} T_j T_j^*
    \Bigr) &\qquad&\text{as~$\kappa_j^*\kappa_i = \delta_{ij}$}.
\end{alignat*}
From this and~$\sum_{j \in I} T_jT_j^* < \infty$,
    it follows that~$(\sum_{j \in S} \kappa_j T_j^*)_S$
    is ultranorm Cauchy.
    Define~$T \equiv ( \sum_{j \in I} \kappa_jT_j^* )^*$,
        where the sum is understood to converge ultranorm.
The assignment~$f \mapsto \kappa_j^*f$
    yields a bounded~$C(B,B)$-linear map between
    the Hilbert~$C(B,B)$-modules~$C(B,A)$ and~$C(B,A_j)$
    and is therefore ultranorm continuous.\cite[\S148]{bas}
    Hence
        $\kappa_j^* T^*
             =  \kappa_j^* \sum_{i} \kappa_i T_i^*
             =  \sum_i \kappa_{j }^*\kappa_i T_i^*
            =  T_j^*$
            and so~$T \kappa_j = T_j$ as desired.
Finally, to show uniqueness,
    assume~$T'\colon A \to B$ is any (other)
    morphism with~$T' \kappa_i = T_i$.
    Note that~$\sum_j \kappa_j \kappa_j^*$
       converges ultrastrongly
       and so ultranorm as well.
With similar reasoning as before,
        the map~$f \mapsto T' f$ is ultranorm continuous,
        hence
\begin{equation*}
    T'
    \ =\ T' \sum_{j \in I} \kappa_j \kappa_j^*
    \ =\ \sum_{j \in I} T' \kappa_j \kappa_j^*
    \ =\ \sum_{j \in I} T_j \kappa_j^*
    \ =\ \sum_{j \in I} T \kappa_j \kappa_j^*
    \ =\ T,
\end{equation*}
    where the sums converge ultranorm.
\end{proof}
